---
title: "Logistic Regression"
format: 
  html:
    fig-width: 12
    fig-height: 9
editor: visual
execute: 
  echo: false
code-tools: true
---

## The Effect of Religion on Personal Support for Same-Sex Marriage

[CSV](https://raw.githubusercontent.com/zief0002/bespectacled-antelope/main/data/same-sex-marriage.csv)

[Codebook](http://zief0002.github.io/bespectacled-antelope/codebooks/same-sex-marriage.html)

### Examining the Relationships

Religious service attendance and type of denomination are thought to be associated with support for or against same-sex marriage. 1746 participants were recruited to take a survey that asked for their denomination, service attendance frequency, and support for or against same-sex marriage (0 = no; 1 = yes). A Likert-scale was used to quantify the service attendance frequency predictor (0 = never; 4 = every week), making it necessary to determine whether this predictor should be treated as continuous or categorical before any regressing can be done. The below figure shows the relationship between support and attendance. Responses are dispersed above and below 0 and 1 on the y-axis to demonstrate that numerous data points were used to create the figure, although all these data points fall on either 0 or 1.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-scatter
#| fig-cap: "Scatterplot with smoother showing the relationship between religious service attendance and support for same-sex marriage."

library(corrr)
library(broom)
#library(educate) 
library(gt)
library(lmtest)
library(patchwork)
library(texreg)
library(stargazer)
library(tidyverse)
library(knitr)
library(kableExtra)
#library(AICcmodavg)
library(performance)

ssm <- read_csv("same-sex-marriage.csv") 
ssm_prop = ssm |> 
  group_by(attendance, support) |> 
  summarize(N = n()) |> 
  mutate(
    Prop = N / sum (N)
  ) 

ggplot(data = ssm, aes(x = attendance, y = jitter(support))) +
  geom_point() +
  geom_smooth(method='loess', se = FALSE) +
  theme_bw() +
  xlab("Religious Service Attendance") +
  scale_y_continuous(
    name = "Support for Same-Sex Marriage",
    breaks = c(0, 0.25, 0.50, 0.75, 1)
  )
```

The relationship between proportion of support and attendance appears to be linear, meaning the Likert-scale data can be treated as if it were continuous.

### Creating and Comparing Models

Five models were created and compared to determine which most appropriately represents the data. The table below shows these models.

```{r}
#| label: tbl-q10
#| tbl-cap: "Table 1. Five candidate models predicting variation in the log-odds of supporting same-sex marriage. The denomination predictor was dummy coded (Catholic, Jewish, and Other), with Protestant as the reference group."
#| message: false
#| echo: false
#| warning: false
#| results: asis

##################################################
### Presenting a Table of Logistic Regression Results
##################################################

# NOTE: If you are using quarto, set your caption in the chunk options with tbl-cap:
# In that case, set caption=NULL in the htmlreg() function below.

prop_ssm = ssm |> 
  group_by(age, support) |> 
  summarize(N = n()) |> 
  mutate(
    Prop = N / sum (N)
  ) |>
  ungroup() |>
  filter(support == "1") |>
  mutate(
    Odds = Prop / (1 - Prop),
    Logits = log(Odds)
  )
#ggplot(data = prop_ssm, aes(x = age, y = Logits)) +
#  geom_point(aes(size = N)) +
#  geom_smooth(aes(weight = N), method = "loess", se = FALSE) +
#  theme_bw() +
 # xlab("Age") +
#  ylab("Log-odds of supporting same-sex marriage")


glm.1_quad = glm(support ~ 1 + attendance + I(attendance^2), data = ssm, family = binomial(link = "logit"))

ssm = ssm |>
  mutate(
    catholic = if_else(denomination == "Catholic", 1, 0),
    jewish = if_else(denomination == "Jewish", 1, 0),
    other = if_else(denomination == "Other", 1, 0)
  )

glm.1_jew = glm(support ~ 1 + catholic + other + jewish, data = ssm, family = binomial(link = "logit"))

glm.3 = glm(support ~ 1 + attendance + I(attendance^2) + catholic + other + jewish, data = ssm, family = binomial(link = "logit"))

glm.5 = glm(support ~ 1 + attendance + I(attendance^2) + catholic + other + jewish + age + friends, data = ssm, family = binomial(link = "logit"))

glm.6 = glm(support ~ 1 + attendance + I(attendance^2) + catholic + other + jewish + age + friends + female, data = ssm, family = binomial(link = "logit"))

#aictab(
 # cand.set = list(glm.1_quad,  glm.3, glm.5, glm.6),
#  modnames = c("Attendance_Q", "Attendance_Q & Denomination", "Age & Friends", "Age, Friends, & Female")
#)

htmlreg(
  l = list(glm.1_quad, glm.1_jew, glm.3, glm.5, glm.6),
  stars = numeric(0),    #No p-value stars
  digits = 2,
  padding = 20,          #Add space around columns (you may need to adjust this via trial-and-error)
  custom.model.names = c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5"), 
  custom.coef.names = c("Intercept", "Attendance_L", "Attendance_Q", 
                        "Catholic", "Jewish", "Other", 
                        "Age", "Friends", "Female"),
  reorder.coef = c(2:9, 1), #Put intercept at bottom of table
  include.aic = FALSE, #Omit AIC
  include.bic = FALSE, #Omit BIC
  include.nobs = FALSE,  #Omit sample size
  include.loglik = FALSE,   #Omit log-likelihood
  custom.gof.rows = list(
    AICc = c(2126.47, 2196.62, 2085.59, 1967.75, 1959.48),
    R2 = (2238.973 - c(2120.452, 2188.596, 2085.592, 1967.755, 1959.48)) /  2238.973),
  reorder.gof = c(3, 1, 2),
  caption = NULL,
  caption.above = TRUE, #Move caption above table
  inner.rules = 1, #Include line rule before model-level output
  outer.rules = 1 , #Include line rules around table
  custom.note = "The $R^2$ value is based on the proportion of reduced deviance from the intercept-only model."
)
```

### Selecting a Model
Model 5 was the most appropriate model for the data. Here is its equation.

$$
\begin{split}
\mathrm{ln(\frac{\hat{\pi_i}}{1+\hat{\pi_i}})} &= -0.19 + 0.14(\mathrm{Attendance}_i) - 0.12(\mathrm{Attendance}_i^2) 

{catholoic, jewish, other, age, friends, female, intercept \
\end{split}
$$ 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-fits
#| fig-cap: "Generalized linear models showing the relationship between religious service attendance and support for same-sex marriage, for male respondents of mean age (47.7 in the dataset) who do not have LGBTQ+ friends."
#tidy(glm.6)
avg_age = ssm |>
  summarize(m = mean(age))
# Plot the fitted equations: Probability of supporting same-sex marriage
ggplot(data = ssm, aes(x = attendance, y = support)) +
  geom_point(alpha = 0) +
  # Protestants
  geom_function(
    fun = function(x) {exp((-.02533145*47.65865) -0.22172950 -0.03491663*(x) -0.07056294*x^2) / (1 + exp(-0.22172950 -0.03491663*(x) -0.07056294*x^2))},
    linetype = "solid",
    aes(color = "Protestant")
  ) + # -.02533145*age 
  # Catholics
    geom_function(
    fun = function(x) {exp( (-.02533145*47.65865) +0.50210485 -0.22172950 -0.03491663*(x) -0.07056294*x^2) / (1 + exp(0.50210485 -0.22172950 -0.03491663*(x) -0.07056294*x^2))},
    linetype = "solid",
    aes(color = "Catholic")
  ) +
   # Jewish
    geom_function(
    fun = function(x) {exp((-.02533145*47.65865) +1.75550397 -0.22172950 -0.03491663*(x) -0.07056294*x^2) / (1 + exp(1.75550397 -0.22172950 -0.03491663*(x) -0.07056294*x^2))},
    linetype = "solid",
    aes(color = "Jewish")
  ) +
   # Other
    geom_function(
    fun = function(x) {exp((-.02533145*47.65865) +0.05528007 -0.22172950 -0.03491663*(x) -0.07056294*x^2) / (1 + exp(0.05528007 -0.22172950 -0.03491663*(x) -0.07056294*x^2))},
    linetype = "solid",
    aes(color = "Other")
  ) +
  theme_light() +
  scale_color_manual(name='Denomination',
                     breaks=c('Protestant', 'Catholic', 'Jewish', 'Other'),
                     values=c('Protestant'='red', 'Catholic'='blue', 'Jewish'='purple', 'Other' = 'green')) +
  xlab("Religious Service Attendance") +
  ylab("Predicted probability of supporting same-sex marriage") +
  ylim(0, 1)
```



